---
title: "Mullen TreeFlux Analysis"
author: "JM"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    code_folding: hide
  word_document:
    toc: true
  pdf_document:
    toc: true
---

Welcome to the TEMPEST Flood Tree Greenhouse Gas Flux dataset analysis!

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(ggplot2)
theme_set(theme_bw())
library(dplyr)
library(lubridate)
library(compasstools)
library(arrow)
library(broom)
library(patchwork)
```

# TEMPEST flood dates

```{r flood-dates}
flood_date <- tribble(~Event, ~Start, ~End,
                      "T1", "2022-06-22", "2022-06-22",
                      "T2", "2023-06-06", "2023-06-07",
                      "T3", "2024-06-11", "2024-06-13")
flood_date$Start <- ymd(flood_date$Start)
flood_date$End <- ymd(flood_date$End)

knitr::kable(flood_date)
```

# Read data

```{r read-data}
# Read in data from Apache Parquet file -----
FLUX_DATA_FILE <- "tempest_tree_ghg_fluxes.parquet"
message("Reading ", FLUX_DATA_FILE)
my_data <- read_parquet(FLUX_DATA_FILE)

code_to_name <- c(ACRU = "Red Maple", FAGR = "Beech", LITU = "Tulip Poplar")

TREE_MAPPING_FILE <- "ancillary_data/sapflow_tree_mapping.csv"
message("Reading ", TREE_MAPPING_FILE)
tree_map <- read.csv(TREE_MAPPING_FILE,
                     stringsAsFactors = FALSE) %>%
  mutate(across(where(is.character), trimws)) %>%
  transmute(ID = Sapflux_ID,
            Species_from_map = unname(code_to_name[Species_Code])) %>%
  distinct(ID, .keep_all = TRUE)

my_data <- my_data %>%
  left_join(tree_map, by = "ID") %>%
  mutate(Species = coalesce(Species, Species_from_map)) %>%
  select(-Species_from_map)

message("Trees with NA species: ", sum(is.na(my_data$Species)))
```

# Prep Soil Temp Data

```{r, eval=FALSE}
read_L2_variable(variable = "soil-temp-5cm", path = "Sensor_Data/", quiet=TRUE)
```

## Take daily avg

```{r}
soil_temp_data <- read_L2_variable(variable = "soil-temp-5cm",
                                   path = "Sensor_Data/", quiet = TRUE)
soil_temp_daily <- soil_temp_data %>%
  mutate(
#    TIMESTAMP = as.POSIXct(TIMESTAMP),
    Date = as.Date(TIMESTAMP),
    Value = as.numeric(Value) ) %>%
  filter(!is.na(Value)) %>%  
  group_by(Site, Plot, Date) %>%
  summarise(
    soil_temp_daily_mean = mean(Value),
    n = n(),
    # TODO: explain this next step (what's 480?)
    coverage = n / 480,      
    .groups = "drop"  )
print(soil_temp_daily)
```

## Graphs of Soil Temp Data (Daily averages over time)

```{r}
ggplot(soil_temp_daily,aes(x = Date, y = soil_temp_daily_mean,
color = Plot, group = Plot)) + geom_line(alpha = 0.7) + facet_wrap(~ Site) +
labs( title = "Daily Mean Soil Temperature by Plot",
   x = "Date",
   y = "Soil Temperature (°C)" ) 
```

## Matching the plot names from tree flux and soil temp

```{r}
soil_temp_daily <- soil_temp_daily %>%
  mutate( Plot = recode(Plot,
         "C" = "Control",
         "F" = "Freshwater",
         "S" = "Seawater" )  )
```

## Joining datasets

```{r}
my_data <- my_data %>%
  left_join(
     soil_temp_daily %>%
    select(Plot, Date, soil_temp_daily_mean),
    by = c("Plot", "Date"))
```

# Numerical summaries -----

print(my_data)
# BBL: print summary of fluxes? A table with mean, min, max and date ranges

```{r}
 my_data %>%
   ggplot(aes( x = Year, y = CH4_lin_flux.estimate, color = Plot )) +
   geom_point(alpha = 0.6) +
   facet_wrap(~ Species) +
   labs( title = "CH4 Flux by Species (2021–2025)" )

my_data %>%
  ggplot(aes( x = Year, y = CO2_lin_flux.estimate, color = Plot )) +
  geom_point(alpha = 0.6) +
  facet_wrap(~ Species) +
  labs( title = "CO2 Flux by Species (2021–2025)", y = "CO2 Flux" )
```



# Numerical summaries

```{r numerical-summaries}
print(my_data)
# BBL: print summary of fluxes? A table with mean, min, max and date ranges
```

# Plots
### Initial visualizations  of Raw Data


## Overall data, by species

```{r overall-by-species}
my_data %>%
  ggplot(aes( x = Year, y = CH4_lin_flux.estimate, color = Plot )) +
  geom_point(alpha = 0.6) +
  facet_wrap(~ Species) +
  labs( title = "CH4 Flux by Species (2021–2025)" )

my_data %>%
  ggplot(aes( x = Year, y = CO2_lin_flux.estimate, color = Plot )) +
  geom_point(alpha = 0.6) +
  facet_wrap(~ Species) +
  labs( title = "CO2 Flux by Species (2021–2025)", y = "CO2 Flux" )
```

## Flux ranges by species

```{r flux-ranges}
# Overall flux range (using boxplot) for each gas, by tree species
ggplot(my_data,aes(x = Species, y = CH4_lin_flux.estimate, fill = Plot)) +
  geom_boxplot() +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

ggplot(my_data, aes(x = Species,  y = CO2_lin_flux.estimate, fill = Plot)) +
  geom_boxplot() +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

## Spaghetti plots of Raw Data

```{r}
ggplot(my_data, aes(x = Date, y = CH4_lin_flux.estimate,
                    group = ID, color = Plot)) +
  geom_line(alpha = 0.2) +
  facet_wrap(~ Species) +
  labs(title = "CH4 Flux Over Time by Species and Plot",
       x = "Date",
       y = "CH4 Flux")

ggplot(my_data, aes(x = Date,
                    y = CO2_lin_flux.estimate,
                    group = ID,
                    color = Plot)) +
  geom_line(alpha = 0.2) +
  facet_wrap(~ Species) +
  labs(
    title = "CO2 Flux Over Time by Species and Plot",
    x = "Date",
    y = "CO2 Flux" )
```

## Heat Maps of of Raw Data

```{r}
heat_data_CH4 <- my_data %>%
  group_by(Species, Year) %>%
  summarize(
    mean_flux = mean(CH4_lin_flux.estimate, na.rm = TRUE),
    .groups = "drop" )
ggplot(heat_data_CH4, aes(x = Year, y = Species, fill = mean_flux)) +
  geom_tile() +
  scale_fill_viridis_c() +
  labs( title = "Mean CH4 Flux by Species and Year", x = "Year", y = "Species",
        fill = "Mean CH4 Flux") + theme_minimal()
seawater_rows <- my_data %>%
  filter(Plot == "Seawater")

heat_data_CO2 <- my_data %>%
  group_by(Species, Year) %>%
  summarize(mean_flux = mean(CO2_lin_flux.estimate, na.rm = TRUE),.groups = "drop")
ggplot(heat_data_CO2,aes(x = Year, y = Species, fill = mean_flux)) +
  geom_tile() +
  scale_fill_viridis_c() +
  labs(title = "Mean CO2 Flux by Species and Year", x = "Year",y = "Species",
       fill = "Mean CO2 Flux" ) +theme_minimal()
```

## Relationship Between CH₄ and CO₂ Flux by Species

```{r}
ggplot(my_data, aes(x = CO2_lin_flux.estimate, y = CH4_lin_flux.estimate, color = Species)) +
  geom_point(alpha = 0.4) +
  geom_smooth(method = "lm", se = FALSE) +
  labs(title = "Relationship Between CO2 and CH4 Flux by Species",x = "CO2 Flux",y = "CH4 Flux",
       color = "Species") +
  theme_minimal()
```


## Quarterly Flux for both GHG (raw data)

```{r}
annual_means_CH4 <- my_data %>%
  group_by(Year, Species, Plot) %>%
  summarize( mean_flux = mean(CH4_lin_flux.estimate, na.rm = TRUE),
             se = sd(CH4_lin_flux.estimate, na.rm = TRUE) / sqrt(n()), .groups = "drop")
ggplot(annual_means_CH4, aes( x = Year, y = mean_flux,color = Plot,group = Plot)) +
  geom_line() +
  geom_point(size = 2) +
  geom_errorbar( aes(  ymin = mean_flux - se,ymax = mean_flux + se  ),width = 0.1 ) +
  facet_wrap(~ Species) +
  labs( title = "Mean CH4 Flux by Species and Plot (2021–2025)",
        x = "Year", y = "Mean CH4 Flux" ) 

annual_means_CO2 <- my_data %>%
  group_by(Year, Species, Plot) %>%
  summarize( mean_flux = mean(CO2_lin_flux.estimate, na.rm = TRUE),
             se = sd(CO2_lin_flux.estimate, na.rm = TRUE) / sqrt(n()), .groups = "drop")
ggplot(annual_means_CO2,aes( x = Year, y = mean_flux, color = Plot, group = Plot  )) +
  geom_line() +
  geom_point(size = 2) +
  geom_errorbar( aes(ymin = mean_flux - se,ymax = mean_flux + se ), width = 0.1) +
  facet_wrap(~ Species) +
  labs( title = "Mean CO2 Flux by Species and Plot (2021–2025)",
        x = "Year",
        y = "Mean CO2 Flux" ) 
```

# Creation of Quarters

```{r}
my_data <- my_data %>%
  mutate(
    Date = as.Date(Date),
    Year = year(Date),
    Quarter = quarter(Date),
    YearQuarter = Year + (Quarter-1)/4)

table(my_data$Quarter)

quarterly_CH4 <- my_data %>%
  mutate(is_S7 = ID == "S7") %>%
  group_by(YearQuarter, Species, Plot, is_S7) %>%
  summarize(mean_flux = mean(CH4_lin_flux.estimate, na.rm = TRUE),
            se = sd(CH4_lin_flux.estimate, na.rm = TRUE) / sqrt(n()), .groups = "drop")
```

# Flood-week averages for reference

```{r}
flood_data <- list()
for(i in seq_len(nrow(flood_date))) {
  message("Isolating data for ", flood_date$Event[i])
  flood_data[[i]] <-
    my_data %>%
    filter(
      Date >= flood_date$Start[i],
      Date <= flood_date$End[i]  ) %>%
    mutate(Event = flood_date$Event[i])}
```
```{r}
flood_CH4 <- bind_rows(flood_data) %>%
  mutate(is_S7 = ID == "S7") %>%
  group_by(Event, YearQuarter, Species, Plot, is_S7) %>%
  summarize(
    mean_flux = mean(CH4_lin_flux.estimate, na.rm = TRUE),
    se = sd(CH4_lin_flux.estimate, na.rm = TRUE) / sqrt(n()),
    .groups = "drop" )
```

# Quarterly Methane Flux with flood averages emphasized

```{r}
ggplot(quarterly_CH4,
       aes(x = YearQuarter, y = mean_flux, color = Plot, group = Plot)) +
  geom_line() +
  geom_point(size = 2) +
  geom_errorbar(
    aes(ymin = mean_flux - se,
        ymax = mean_flux + se),
    width = 0.05) +
  # Flood overlay (T1, T2, T3 labeled in legend)
  geom_point(
    data = flood_CH4,
    aes(x = YearQuarter,
        y = mean_flux,
        shape = Event),
    color = "black",
    size = 1.5,
    inherit.aes = FALSE) +
  facet_wrap(~ Species) +
  labs( title = "Quarterly CH4 Flux with Flood Events Highlighted",
        x = "Year-Quarter",
        y = "CH4 Flux",
        shape = "Flood event") +
  scale_shape_manual(values = c(
    T1 = 17,
    T2 = 17,
    T3 = 17)) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

```

# Isolating Red Maple Trees (finding outlier (S7))

```{r}
my_data %>%
  filter(Species == "Red Maple") %>%
  ggplot(aes(x = Date, y = CH4_lin_flux.estimate)) +
  geom_line(alpha = 0.7) +
  facet_wrap(~ ID) +
  labs(title = "Red Maple CH4 Flux by Individual Tree")
my_data %>%
  filter(Species == "Red Maple") %>%
  ggplot(aes(x = Date, y = CO2_lin_flux.estimate)) +
  geom_line(alpha = 0.7) +
  facet_wrap(~ ID) 
```

# Separate S7 from all other trees

```{r}
s7_CH4 <- my_data %>%
  filter(ID == "S7") %>%
  group_by(YearQuarter, Species, Plot) %>%
  summarize(mean_flux = mean(CH4_lin_flux.estimate, na.rm = TRUE),
            se = sd(CH4_lin_flux.estimate, na.rm = TRUE) / sqrt(n()), .groups = "drop"  )
quarterly_CH4_nos7 <- my_data %>%
  filter(ID != "S7") %>%
  group_by(YearQuarter, Species, Plot) %>%
  summarize(mean_flux = mean(CH4_lin_flux.estimate, na.rm = TRUE),
            se = sd(CH4_lin_flux.estimate, na.rm = TRUE) / sqrt(n()), .groups = "drop" )
quarterly_CH4_S7 <- s7_CH4
flood_points <- flood_date %>%
  mutate(YearQuarter = year(Start) + (quarter(Start)-1)/4) %>%
  select(Event, YearQuarter)
flood_S7_CH4 <- s7_CH4 %>%
  inner_join(flood_points, by = "YearQuarter")
```

#Quarterly CH4 Flux (S7 excluded from means, shown separately)

```{r}
ggplot() +
  # OTHER TREES (baseline)
  geom_line(
    data = quarterly_CH4_nos7,
    aes(x = YearQuarter, y = mean_flux, color = Plot, group = Plot)) +
  geom_point(
    data = quarterly_CH4_nos7,
    aes(x = YearQuarter, y = mean_flux, color = Plot)) +
  # S7 (highlighted)
  geom_line(
    data = s7_CH4,
    aes(x = YearQuarter, y = mean_flux),
    color = "red",
    linewidth = 1 ) +
  geom_point(data = s7_CH4,
             aes(x = YearQuarter, y = mean_flux),
             color = "red", size = 2) +
  # FLOOD EVENTS ON S7 
  geom_point(
    data = s7_CH4 %>%
      inner_join(flood_points, by = "YearQuarter"),
    aes(x = YearQuarter, y = mean_flux, shape = Event),
    color = "black",
    size = 3 ) +
  facet_wrap(~ Species) +
  labs( title = "Quarterly CH4 Flux (S7 excluded from means, shown separately)",
        x = "Year-Quarter",
        y = "CH4 Flux",
        shape = "Flood event") +
  scale_shape_manual(values = c(
    T1 = 17,
    T2 = 17,
    T3 = 17 )) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

#S7 only 

```{r}
ggplot(quarterly_CH4_S7, aes(x = YearQuarter, y = mean_flux)) +
  geom_line(color = "red", linewidth = 1) + geom_point(color = "red", size = 2) +
  geom_errorbar(aes(ymin = mean_flux - se, ymax = mean_flux + se),
                width = 0.05, color = "red" ) +
  geom_point(
    data = flood_S7_CH4,
    aes(x = YearQuarter, y = mean_flux),
    color = "black",
    size = 2  ) +
  labs( title = "S7 Quarterly CH4 Flux with Flood Periods Highlighted",
        x = "Year-Quarter",
        y = "CH4 Flux") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

#S7 Excluded from CH4 Flux

```{r}
flood_plot_data <- quarterly_CH4_nos7 %>%
  inner_join(flood_points, by = "YearQuarter")
ggplot() +  geom_line(
  data = quarterly_CH4_nos7,
  aes(x = YearQuarter, y = mean_flux, color = Plot, group = Plot) ) +
  geom_point(
    data = quarterly_CH4_nos7,aes(x = YearQuarter, y = mean_flux, color = Plot) ) +
  geom_point(data = flood_plot_data,
             aes(x = YearQuarter, y = mean_flux, shape = Event),
             color = "black",
             size = 3 ) +
  facet_wrap(~ Species) +
  labs(title = "Quarterly CH4 Flux with Flood Events Across Treatments",
       x = "Year-Quarter",
       y = "CH4 Flux",
       shape = "Flood event" ) +
  scale_shape_manual(values = c(T1 = 17, T2 = 17, T3 = 17)) +
  theme_minimal() + theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

# Visualizing Soil Temp Data x GHG Tree Flux

## GHG Tree Flux x Soil Temp (colored by species)

```{r}
ggplot(my_data, aes(x = soil_temp_daily_mean,
             y = CH4_lin_flux.estimate,
             color = Species)) +
 geom_point(alpha = 0.5) +
 geom_smooth(method = "lm", se = FALSE) +
 labs(
     x = "Daily Mean Soil Temperature (°C)",
     y = "CH₄ Flux" )

ggplot(my_data,aes(x = soil_temp_daily_mean,
                   y = CO2_lin_flux.estimate,
                   color = Species)) +
geom_point(alpha = 0.5) + geom_smooth(method = "lm", se = FALSE)

```

## CH4 Tree Flux x Soil Temp (colored by species)

```{r}
ggplot(my_data, aes(x = soil_temp_daily_mean,
             y = CH4_lin_flux.estimate,
             color = Species)) +
 geom_point(alpha = 0.5) +
 geom_smooth(method = "lm", se = FALSE) +
 labs(
     x = "Daily Mean Soil Temperature (°C)",
     y = "CH4 Flux" )
```


## CO2 Tree Flux x Soil Temp (colored by species)

```{r}
ggplot(my_data,aes(x = soil_temp_daily_mean,
                   y = CO2_lin_flux.estimate,
                   color = Species)) +
geom_point(alpha = 0.5) + geom_smooth(method = "lm", se = FALSE)
```


## Methane and Soil Temp Over Time 

```{r}
ggplot(my_data, aes(x = Date)) +
 geom_line(aes(y = CH4_lin_flux.estimate, color = "CH4"), alpha = 0.5) +
geom_line(aes(y = soil_temp_daily_mean, color = "Soil Temp"), alpha = 0.7) +
 facet_wrap(~ Species) +
 scale_color_manual(values = c("CH4" = "darkgreen", "Soil Temp" = "red")) +
labs( title = "CH4 Flux and Soil Temperature Over Time",
     y = "Value (different scales)",
   color = "Variable"    )
```
# Carbon Dioxide and Soil Temp Over Time
```{r}
ggplot(my_data, aes(x = Date)) +
 geom_line(aes(y = CO2_lin_flux.estimate, color = "C02"), alpha = 0.5) +
geom_line(aes(y = soil_temp_daily_mean, color = "Soil Temp"), alpha = 0.7) +
 facet_wrap(~ Species) +
 scale_color_manual(values = c("CO2" = "darkgreen", "Soil Temp" = "red")) +
labs( title = "CO2 Flux and Soil Temperature Over Time",
     y = "Value (different scales)",
   color = "Variable"    )
```

   
## Methane vs Soil Temp (by plot)

```{r}
ggplot(my_data, aes(x = soil_temp_daily_mean, y = CH4_lin_flux.estimate)) +
   geom_point(aes(color = Plot), alpha = 0.5) +
       geom_smooth(method = "lm", se = FALSE) +
     facet_wrap(~ Species) +
    geom_vline(data = flood_date,
aes(xintercept = NA), linetype = "dashed")  # placeholder style

```
## CO2 Vs Soil Temp (by plot)
```{r}
ggplot(my_data, aes(x = soil_temp_daily_mean, y = CO2_lin_flux.estimate)) +
   geom_point(aes(color = Plot), alpha = 0.5) +
       geom_smooth(method = "lm", se = FALSE) +
     facet_wrap(~ Species) +
    geom_vline(data = flood_date,
aes(xintercept = NA), linetype = "dashed")  # placeholder style
```

## Soil Temperature × Treatment Effects on CH4 Flux

```{r}
ggplot(my_data, aes(x = soil_temp_daily_mean,
                  y = CH4_lin_flux.estimate,
                     color = Plot)) +
     geom_point(alpha = 0.4) +
     geom_smooth(method = "lm", se = FALSE) +
     facet_wrap(~ Species) +
     labs(title = "Soil Temperature × Treatment Effects on CH4 Flux")
```
## Soil Temp x Treatment Effects on CO2 Flux
```{r}
ggplot(my_data, aes(x = soil_temp_daily_mean,
                  y = CO2_lin_flux.estimate,
                     color = Plot)) +
     geom_point(alpha = 0.4) +
     geom_smooth(method = "lm", se = FALSE) +
     facet_wrap(~ Species) +
     labs(title = "Soil Temperature × Treatment Effects on CO2 Flux")
```


## Nonlinear Temperature Response of CH4 (What is the actual shape of the temperature–methane relationship?)

### Locally Estimated Scatterplot Smoothing = LOESS

```{r}
ggplot(my_data, aes(soil_temp_daily_mean, CH4_lin_flux.estimate)) +
  geom_point(alpha = 0.4, aes(color = Species)) +
  geom_smooth(method = "loess", se = FALSE) +
  facet_wrap(~ Plot) +
  labs(title = "Nonlinear Temperature Response of CH4")
```
## Nonlinear Temperature Response of CO2 (What is the actual shape of the temperature–carbon dioxide relationship?)
```{r}
ggplot(my_data, aes(soil_temp_daily_mean, CO2_lin_flux.estimate)) +
  geom_point(alpha = 0.4, aes(color = Species)) +
  geom_smooth(method = "loess", se = FALSE) +
  facet_wrap(~ Plot) +
  labs(title = "Nonlinear Temperature Response of CO2")
```


# Species Differences in Temperature Sensitivity (Do different species have different average rates of response to temperature?)

```{r}
ggplot(my_data, aes(soil_temp_daily_mean,
                    CH4_lin_flux.estimate,
                    color = Species)) +
  geom_point(alpha = 0.4) +
  geom_smooth(method = "lm", se = FALSE) +
  labs(title = "Species Differences in Temperature Sensitivity (CH4)")
```
```{r}
ggplot(my_data, aes(soil_temp_daily_mean,
                    CO2_lin_flux.estimate,
                    color = Species)) +
  geom_point(alpha = 0.4) +
  geom_smooth(method = "lm", se = FALSE) +
  labs(title = "Species Differences in Temperature Sensitivity (CO2)")
```

## Relationship between Soil Temp and Methane Flux
```{r}
 ggplot(my_data,
    aes(x = soil_temp_daily_mean,
            y = CH4_lin_flux.estimate,
            color = Species)) +
     geom_point(alpha = 0.35) +
     geom_smooth(method = "loess", se = TRUE) +
     labs(
         x = "Daily Mean Soil Temperature (°C)",
         y = "CH4 Flux",
         title = "Relationship Between Soil Temperature and CH4 Flux" ) 
```

## Relationship between Soil Temp and CO2 Flux
```{r}
ggplot(my_data,
       aes(x = soil_temp_daily_mean,
            y = CO2_lin_flux.estimate,
            color = Plot)) +
     geom_point(alpha = 0.4) +
     geom_smooth(method = "loess", se = FALSE) +
     facet_wrap(~Species) +
     labs(
         title = "Relationship Between Daily Mean Soil Temperature and CO2 Flux",
         x = "Daily Mean Soil Temperature (°C)",
         y = expression(CO[2]~"Flux"),
         color = "Treatment")
```


## Facet by species, color by plot
```{r}
ggplot(my_data,
 aes(x = soil_temp_daily_mean,
       y = CH4_lin_flux.estimate,
            color = Plot)) +
     geom_point(alpha = 0.35) +
     geom_smooth(method = "loess", se = FALSE) +
     facet_wrap(~Species)
```
```{r}
ggplot(my_data,
 aes(x = soil_temp_daily_mean,
       y = CO2_lin_flux.estimate,
            color = Plot)) +
     geom_point(alpha = 0.35) +
     geom_smooth(method = "loess", se = FALSE) +
     facet_wrap(~Species)
```

## Facet by Plot, color by Species
```{r}
ggplot(my_data,
  aes(x = soil_temp_daily_mean,
            y = CH4_lin_flux.estimate,
            color = Species)) +
     geom_point(alpha = 0.35) +
     geom_smooth(method = "loess", se = FALSE) +
     facet_wrap(~Plot)
```
```{r}
ggplot(my_data,
  aes(x = soil_temp_daily_mean,
            y = CO2_lin_flux.estimate,
            color = Species)) +
     geom_point(alpha = 0.35) +
     geom_smooth(method = "loess", se = FALSE) +
     facet_wrap(~Plot)
```

## Seasonal and Flood-Driven Dynamics of Soil Temperature and Greenhouse Gas Fluxes (CH4 and CO2)
```{r}

# CO2 quarterly summary
quarterly_CO2 <- my_data %>%
  group_by(YearQuarter, Plot, Species) %>%
  summarize(
    mean_flux = mean(CO2_lin_flux.estimate, na.rm = TRUE),
    se = sd(CO2_lin_flux.estimate, na.rm = TRUE) / sqrt(n()),
    .groups = "drop" )

# Soil quarterly summary
quarterly_soil <- my_data %>%
  group_by(YearQuarter, Plot, Species) %>%
  summarize(
    mean_soil = mean(soil_temp_daily_mean, na.rm = TRUE),
    .groups = "drop")

# Create flood dataset for soil temperature
# -------------------------
flood_soil <- quarterly_soil %>%
  inner_join(
    flood_date %>%
      mutate(YearQuarter = year(Start) + (quarter(Start) - 1) / 4),
    by = "YearQuarter"
  ) %>%
  select(YearQuarter, Plot, Species, mean_soil, Event)

# CH4 plot
p1 <- ggplot(quarterly_CH4,
             aes(x = YearQuarter,
                 y = mean_flux,
                 color = Plot,
                 group = Plot)) +
  geom_line() +
  geom_point(size = 2) +
  geom_errorbar(aes(ymin = mean_flux - se,
                    ymax = mean_flux + se),
                width = 0.05) +
  geom_point(data = flood_CH4,
             aes(x = YearQuarter,
                 y = mean_flux,
                 shape = Event),
             color = "black",
             size = 1.5,
             inherit.aes = FALSE) +
  facet_wrap(~Species) +
  labs(y = "CH₄ Flux",
       x = NULL,
       shape = "Flood event") 

# CO2 plot
p2 <- ggplot(quarterly_CO2,
             aes(x = YearQuarter,
                 y = mean_flux,
                 color = Plot,
                 group = Plot)) +
  geom_line() +
  geom_point(size = 2) +
  geom_errorbar(aes(ymin = mean_flux - se,
                    ymax = mean_flux + se),
                width = 0.05) +
  geom_point(data = flood_CH4,
             aes(x = YearQuarter,
                 y = mean_flux,
                 shape = Event),
             color = "black",
             size = 1.5,
             inherit.aes = FALSE) +
  facet_wrap(~Species) +
  labs(y = "CO₂ Flux",
       x = NULL,
       shape = "Flood event") 

# Soil temperature plot
p3 <- ggplot(quarterly_soil,
             aes(x = YearQuarter,
                 y = mean_soil,
                 color = Plot,
                 group = Plot)) +
  geom_line(linewidth = 1) +
  geom_point(data = flood_soil,
             aes(x = YearQuarter,
                 y = mean_soil,
                 shape = Event),
             color = "black",
             size = 1.5,
             inherit.aes = FALSE) +
  facet_wrap(~Species) +
  labs(y = "Soil Temperature (°C)",
       x = "Year-Quarter",
       shape = "Flood event") 

# Combine all plots
p1 / p2 / p3 +
  plot_layout(ncol = 1, guides = "collect") &
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

# All Trees GHG Relationship
```{r}
ggplot(my_data,
     aes(x = soil_temp_daily_mean,
            y = CH4_lin_flux.estimate)) +
     geom_point(alpha = 0.4) +
     geom_smooth(method = "loess", se = FALSE, color = "blue") +
     facet_wrap(~ID, scales = "free") +
     labs(
         title = "Methane Flux vs Soil Temperature by Tree",
         x = "Daily Mean Soil Temperature (°C)",
         y = "CH4 Flux") 
```

```{r}
ggplot(my_data,
       aes(x = soil_temp_daily_mean,
            y = CO2_lin_flux.estimate)) +
     geom_point(alpha = 0.4) +
     geom_smooth(method = "loess", se = FALSE, color = "blue") +
     facet_wrap(~ID, scales = "free") +
     labs(
         title = "Carbon Dioxide Flux vs Soil Temperature by Tree",
         x = "Daily Mean Soil Temperature (°C)",
         y = "CO2 Flux")
```

# Methane x Soil by Plot

## Methane Flux vs. Daily Mean Soil Temperature (Control Trees)
```{r}
control_data <- my_data %>%
   filter(Plot == "Control")
 
 ggplot(control_data,
        aes(x = soil_temp_daily_mean,
            y = CH4_lin_flux.estimate)) +
     geom_point(alpha = 0.4) +
     geom_smooth(method = "loess", se = FALSE, color = "green") +
     facet_wrap(~ID, scales = "free") +
     labs(
         title = "Methane Flux vs. Daily Mean Soil Temperature (Control Trees)",
         x = "Daily Mean Soil Temperature (°C)",
         y = "CH4 Flux")
```

## Methane Flux vs. Daily Mean Soil Temperature (Freshwater Trees)
```{r}
Freshwater_data <- my_data %>%
     filter(Plot == "Freshwater")
 ggplot(Freshwater_data,
        aes(x = soil_temp_daily_mean,
            y = CH4_lin_flux.estimate)) +
     geom_point(alpha = 0.4) +
     geom_smooth(method = "loess", se = FALSE, color = "blue") +
     facet_wrap(~ID, scales = "free") +
     labs(
         title = "Methane Flux vs. Daily Mean Soil Temperature (Freshwater Trees)",
         x = "Daily Mean Soil Temperature (°C)",
         y = "CH4 Flux")
```

## Methane Flux vs. Daily Mean Soil Temperature (Seawater Trees)
```{r}
Seawater_data <- my_data %>%
    filter(Plot == "Seawater")
 ggplot(Seawater_data,
        aes(x = soil_temp_daily_mean,
            y = CH4_lin_flux.estimate)) +
     geom_point(alpha = 0.4) +
     geom_smooth(method = "loess", se = FALSE, color = "red") +
     facet_wrap(~ID, scales = "free") +
     labs(
         title = "Methane Flux vs. Daily Mean Soil Temperature (Seawater Trees)",
         x = "Daily Mean Soil Temperature (°C)",
         y = "CH4 Flux")
```

# Carbon Dioxide x Soil by Plot

## CO2 Flux vs. Daily Mean Soil Temperature (Control Trees)
```{r}
control_data <- my_data %>%
    filter(Plot == "Control")
ggplot(control_data,
       aes(x = soil_temp_daily_mean,
           y = CO2_lin_flux.estimate)) +
    geom_point(alpha = 0.4) +
    geom_smooth(method = "loess", se = FALSE, color = "green") +
    facet_wrap(~ID, scales = "free") +
    labs(
        title = "CO2 Flux vs. Daily Mean Soil Temperature (Control Trees)",
        x = "Daily Mean Soil Temperature (°C)",
        y = "CO2 Flux")
```

## CO2 Flux vs. Daily Mean Soil Temperature (Freshwater Trees)
```{r}
Freshwater_data <- my_data %>%
    filter(Plot == "Freshwater")
ggplot(Freshwater_data,
       aes(x = soil_temp_daily_mean,
           y = CO2_lin_flux.estimate)) +
    geom_point(alpha = 0.4) +
    geom_smooth(method = "loess", se = FALSE, color = "blue") +
    facet_wrap(~ID, scales = "free") +
    labs(
        title = "CO2 Flux vs. Daily Mean Soil Temperature (Freshwater Trees)",
        x = "Daily Mean Soil Temperature (°C)",
        y = "CO2 Flux")
```

## CO2 Flux vs. Daily Mean Soil Temperature (Seawater Trees)
```{r}
Seawater_data <- my_data %>%
    filter(Plot == "Seawater")
ggplot(Seawater_data,
       aes(x = soil_temp_daily_mean,
           y = CO2_lin_flux.estimate)) +
    geom_point(alpha = 0.4) +
    geom_smooth(method = "loess", se = FALSE, color = "red") +
    facet_wrap(~ID, scales = "free") +
    labs(
        title = "CO2 Flux vs. Daily Mean Soil Temperature (Seawater Trees)",
        x = "Daily Mean Soil Temperature (°C)",
        y = "CO2 Flux")
```

# Carbon Dioxide x Soil by Plot

## CO2 Flux vs. Soil Temperature (Red Maple Trees)
```{r}
 red_maple_data <- my_data %>%
  filter(Species == "Red Maple")
 ggplot(red_maple_data,
        aes(x = soil_temp_daily_mean,
            y = CO2_lin_flux.estimate)) +
     geom_point(alpha = 0.4) +
     geom_smooth(method = "loess", se = FALSE, color = "red") +
     facet_wrap(~ID, scales = "free") +
     labs(
         title = "CO2 Flux vs Soil Temperature (Red Maple Trees)",
       x = "Soil Temperature (°C)",
         y = "CO2 Flux")
```

## CO2 Flux vs. Soil Temperature (Beech Trees)
```{r}
beech_data <- my_data %>%
    filter(Species == "Beech")
 ggplot(beech_data,
        aes(x = soil_temp_daily_mean,
            y = CO2_lin_flux.estimate)) +
     geom_point(alpha = 0.4) +
     geom_smooth(method = "loess", se = FALSE, color = "green") +
     facet_wrap(~ID, scales = "free") +
     labs(
         title = "CO2 Flux vs Soil Temperature (Beech Trees)",
         x = "Soil Temperature (°C)",
         y = "CO2 Flux")
```

## CO2 Flux vs. Soil Temperature (Tulip Poplar Trees)
```{r}
tulip_poplar_data <- my_data %>%
    filter(Species == "Tulip Poplar")
 ggplot(tulip_poplar_data,
        aes(x = soil_temp_daily_mean,
            y = CO2_lin_flux.estimate)) +
     geom_point(alpha = 0.4) +
     geom_smooth(method = "loess", se = FALSE, color = "blue") +
     facet_wrap(~ID, scales = "free") +
     labs(
         title = "CO2 Flux vs Soil Temperature (Tulip Poplar Trees)",
         x = "Soil Temperature (°C)",
         y = "CO2 Flux")
```

# Methane x Soil by Plot

## CH4 Flux vs. Soil Temperature (Red Maple Trees)
```{r}
red_maple_data <- my_data %>%
    filter(Species == "Red Maple")
ggplot(red_maple_data,
       aes(x = soil_temp_daily_mean,
           y = CH4_lin_flux.estimate)) +
    geom_point(alpha = 0.4) +
    geom_smooth(method = "loess", se = FALSE, color = "red") +
    facet_wrap(~ID, scales = "free") +
    labs(
        title = "Methane Flux vs Soil Temperature (Red Maple Trees)",
        x = "Soil Temperature (°C)",
        y = "CH4 Flux")
```

## CH4 Flux vs. Soil Temperature (Beech Trees)
```{r}
beech_data <- my_data %>%
    filter(Species == "Beech")
ggplot(beech_data,
       aes(x = soil_temp_daily_mean,
           y = CH4_lin_flux.estimate)) +
    geom_point(alpha = 0.4) +
    geom_smooth(method = "loess", se = FALSE, color = "green") +
    facet_wrap(~ID, scales = "free") +
    labs(
        title = "Methane Flux vs Soil Temperature (Beech Trees)",
        x = "Soil Temperature (°C)",
        y = "CH4 Flux")
```

## CH4 Flux vs. Soil Temperature (Tulip Poplar Trees)
```{r}
tulip_poplar_data <- my_data %>%
  filter(Species == "Tulip Poplar")
ggplot(tulip_poplar_data,
       aes(x = soil_temp_daily_mean,
           y = CH4_lin_flux.estimate)) +
  geom_point(alpha = 0.4) +
  geom_smooth(method = "loess", se = FALSE, color = "blue") +
  facet_wrap(~ID, scales = "free") +
  labs(
    title = "CH₄ Flux vs Soil Temperature (Tulip Poplar Trees)",
    x = "Soil Temperature (°C)",
    y = "CH₄ Flux" )
```
# Categorize trees' Temp sensitivity 

```{r}
ch4_sensitivity <- my_data %>%
    ungroup() %>%
     reframe(
         lm(CH4_lin_flux.estimate ~ soil_temp_daily_mean) %>% tidy(),
         .by = ID )
ch4_sensitivity
```
```{r}
ch4_sensitivity <- ch4_sensitivity %>%
    filter(term == "soil_temp_daily_mean") %>%
    mutate(
        sensitivity = case_when(
            p.value < 0.05 & estimate > 0 ~ "Positive",
            p.value < 0.05 & estimate < 0 ~ "Negative",
            TRUE ~ "Neutral"
        ))
ch4_sensitivity
```

 # Visualization of Temp Sensitivity
```{r}
ggplot(subset(ch4_sensitivity, sensitivity == "Positive"),
       aes(x = reorder(ID, estimate),
           y = estimate)) +
  geom_col(fill = "darkgreen") +
  coord_flip() +
  labs(title = "Positive Temperature Sensitivity (CH4)",
       x = "Tree ID",
       y = "Slope")
```
 
```{r}
ggplot(subset(ch4_sensitivity, sensitivity == "Neutral"),
       aes(x = reorder(ID, estimate),
           y = estimate)) +
  geom_col(fill = "blue") +
  coord_flip() +
  labs(title = "Neutral Temperature Sensitivity (CH4)",
       x = "Tree ID",
       y = "Slope") 
```

## Checking for negatives 
```{r}
table(ch4_sensitivity$sensitivity)
```

```{r}
negative_data <- ch4_sensitivity %>%
    filter(sensitivity == "Negative")
negative_data <- ch4_sensitivity %>%
    filter(sensitivity == "Negative")
negative_data
```

